Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-13 nips_2018.tex
writing

\title{Learning Reusable Components for Neurally--Guided Bayesian Program Learning}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it represents the linear function $-x+2$ with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{l}{0.4\textwidth}\centering
\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(* real (+ x real))}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* x (+ real x))} \\
  \\\bottomrule
\end{tabular}
\caption{Some learned subroutines for symbolic regression. The system starts with addition, multiplication, and real numbers, and learns to build polynomials up to 4th order.}\label{regressionDSL}
\end{wrapfigure}

@@

How well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, our model on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
our system solves 32/108 of the problems with an average search time of 11 minutes.
After learning,
our system solves 80/108, and does so much faster,
solving them in an average of 29 seconds.

@@

\subsection{List Functions}\label{listSection}
Synthesizing programs that manipulate data structures is
a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of
learning functions that manipulate lists.
We created 225 Lisp-style list manipulation tasks,
each with 15 input/output examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions:
many of the functions are very complicated,
and the agent must learn to solve these complicated problems from only 225 tasks.

@@

We evaluate \system starting from two different initial DSLs, \emph{base} and
\emph{rich}. The base DSL starts with only
low-level primitives such as \code{mapi}, \code{reducei}, and
\code{if}, whereas the rich DSL includes some common structure that can be
built from these such as \code{all}, \code{filter}, \code{slice}, and
\code{index}.
Both are capable of solving all tasks supplied by our dataset.
See Supplement for details on these DSLs.

We found  this domain  difficult to tackle without starting from 
the rich DSL. Our
algorithm, like human learners, requires a spectrum of problems ranging from
easy to hard.
When starting with the base DSL,
we found that there were not enough steppingstones in the curriculum to
get \system off the ground.

@@

  \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
  \% solved &
  \textbf{84\%} &76\%& 63\%&62\%&62\%\\
  Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s

@@

  \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
  \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
  Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s

  \\\midrule\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
  \\\bottomrule
  \end{tabular}
\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved
  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  MDL:
  $-\expect\left[\probability[x|p] \right]$. For domains other than
  symbolic regression MDL is 0 nats. For symbolic regression MDL is a
  proxy for \# of continuous parameters.}\vspace{-0.5cm}\label{baselineComparisons} \end{table}

>>>>

\title{Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{l}{0.4\textwidth}\centering
\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(* real (+ x real))}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* x (+ real x))} \\
  $f_2($\code{x}$)\,=\,$\code{(/ (/ x real) x)}
  \\\bottomrule
\end{tabular}
\caption{Some learned subroutines for symbolic regression. The system starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
\end{wrapfigure}

@@

How well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 32/108 of the problems with an average search time of 11 minutes.
After learning,
it solves 80/108, and does so much faster,
solving them in an average of 29 seconds.
As of the 2017 SyGuS competition,
the best-performing algorithm solves 86/108 problems,
and does so with a
\emph{different hand-engineered DSL for each problem}.
Here  we learned a single DSL
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.

@@

\subsection{List Functions}\label{listSection}
Synthesizing programs that manipulate data structures is
a widely studied problem in the programming languages community~\cite{feser2015synthesizing}.
%with applications to computer aided programming~\cite{solar2008program}.
We consider this problem within the context of
learning functions that manipulate lists.
We created 225 Lisp-style list manipulation tasks,
each with 15 input/output examples (Tbl.~\ref{listexamples}).
Our data set is challenging along two dimensions:
many of the functions are very complicated,
and the agent must learn to solve these complicated problems from only 225 tasks.
Our data set primarily consists of arithmetic operations upon sequences,
and so, in addition to the Lisp primitives provided for the text editing experiments, we additionally start the system out with the following primitives:
\code{mod}, \code{*}, \code{>}, \code{is-square}, \code{is-prime}.

A complete repertoire of higher-order functions is a staple of functional programming standard libraries.
Although we provided \system with some of the standard higher-order functions,
like \code{fold} and \code{unfold},
we did not include others.
When trained on these list functions,
our system rediscovers
and then reuses the higher-order function \code{filter} \textbf{Lucas: not sure if it actually does this! It would be cool if we can write something like this though.}

@@

%% We evaluate \system starting from two different initial DSLs, \emph{base} and
%% \emph{rich}. The base DSL starts with only
%% low-level primitives such as \code{mapi}, \code{reducei}, and
%% \code{if}, whereas the rich DSL includes some common structure that can be
%% built from these such as \code{all}, \code{filter}, \code{slice}, and
%% \code{index}.
%% Both are capable of solving all tasks supplied by our dataset.
%% See Supplement for details on these DSLs.

%% We found  this domain  difficult to tackle without starting from 
%% the rich DSL. Our
%% algorithm, like human learners, requires a spectrum of problems ranging from
%% easy to hard.
%% When starting with the base DSL,
%% we found that there were not enough steppingstones in the curriculum to
%% get \system off the ground.

@@

  %% \\\midrule\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
  %% \% solved &
  %% \textbf{84\%} &76\%& 63\%&62\%&62\%\\
  %% Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s

@@

  %% \\\midrule\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
  %% \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
  %% Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s

  \\\midrule\multicolumn{6}{c}{\emph{List functions}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
  \\\bottomrule
  \end{tabular}
\caption{\% solved w/ 5 sec timeout. Solve time: averaged over solved
  tasks.  RF/DC: trained like RobustFill/DeepCoder.  PCFG: model w/o
  structure learning.  Enum: model w/o any learning.  MDL:
  $-\expect\left[\probability[x|p] \right]$. For domains other than
  symbolic regression MDL is 0 nats. For symbolic regression MDL is a
  proxy for \# of continuous parameters.}\vspace{-0.5cm}\label{baselineComparisons} \end{table}
